---
title: "STTARS_Simulation_4"
author: "M Wilson"
date: "14/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(naniar)

```

# Aims

- Thoughts on how to analyse the second randomisation step of the study.



# Background:

Following the initial randomisation step on admission to either IMV or CPAP, clinicians receive a nudge for each patient in the study, who is on CPAP at each 12 hour handover point.

The nudge asks if there is clinical equipoise around the decision of whether to continue CPAP or start IMV during the next 12 hour period.  

In most cases it is anticipated that the clinicians will not have equipoise around the decision and will continue as they see best.  However, in some circumstances they may not know what the best course of action is.  Given there is little evidence to inform this decision, they may wish to learn from the results of whatever decision is made.  In this case, they would be encouraged to submit the patient in question for repeat randomisation to either starting IMV or continuing current CPAP.  

Having obtained the allocation decision, the patient proceeds to either a further 12 hour period of CPAP, or to intubation and starting of IMV.  



# Reverse Thinking

Start from the perspective of analysing an observational study with the aim of identifying the optimal duration of cpap treatment. 

You would have a dataset containing individual patients, their respiratory support status at each 12 hour period from the start of the study and an indication of whether their status changed during the subsequent 12 hour period of observation.  

You might use an instrumental variable approach to answer this question, where the outcome (Y) is continuous (for simplicity initially) such as length of stay on ICU (days).  The exposure (X) could be continuous e.g. number of hours on CPAP, or binary e.g. CPAP at each 12 hour time point Y|N.  

In our observational study, the instrument (Z) would be the physician's preference for either continuing CPAP (Z = 0) or starting IMV (Z = 1) at each 12 hour time point.  In our proposed RCT, we capture the physician preference in the form of an accept/reject random nudge to either Z = 0 | Z = 1. 

Back to the observational study, we must demonstrate the assumptions of the IV analysis, namely, 
that the instrument Z is not associated with unobserved confounders (U) which determine the values of both Y (length of ICU stay) and X (duration of CPAP, hours).  

One example of an unmeasured confounder in this case might be ACE genotype which might predispose to more severe infection, producing longer length of stay and shorter duration of CPAP?

The instrument (Z) must be correlated with the exposure (X) but not with outcome (Y) other than through mediating X.  A strong instrument is one which causes large variation in X and this relationship must be estimated.  A weak instrument will have a low power to detect a causal effect and may induce bias.  


## Causal Assumptions

The goal is to identify and estimate the causal effect of duration of CPAP (X) on length of ICU stay (Y).  There are likely to be many common causes of both X and Y and these will not all be observable.  Therefore, we seek to use physician preference for continuing cpap over starting IMV as an instrument for X.  If the physician has a preference for IMV then the duration of CPAP will be shorter, but if the physician thinks there is benefit from continued application of CPAP then they may opt to continue for a longer duration. 

We assume that the instrument is only able to directly influence Y through it's action on Z.  


```{r}
# let's say you have 100 patients on cpap at the beginning:

n <- c(1:100)

# each of these patients has a length of stay on ICU (hours) between 0 and 100 days

los <- 
  runif(100, min = 0, max = 2400)

# every patient received cpap, but for varying time periods (hours), between 1 and 120 hours

cpap_duration <- 
  runif(100, min = 1, max = 120)


# combine into a basic dataframe:

tbl1 <- 
  tibble(
    n = n,
    los = los,
  )

# make sure you can't be on cpap longer than your los:

tbl1 <- 
  tbl1 %>% 
  mutate(cpap = runif(100, min = 1, max = los))

# check the duration of cpap does not exceed los:

tbl1 %>% 
  filter(cpap > los) # 0

# because we are only interested in the first 5 days of treatment (as this is the randomisation window and assuming most people on cpap will be below this duration), cap the max cpap value to 120:

tbl1 <- 
  tbl1 %>% 
  mutate(cpap = if_else(cpap > 120, 120, cpap))

# floor each to a round hour

tbl1 <- 
  tbl1 %>% 
  mutate(los = floor(los)) %>% 
  mutate(cpap = floor(cpap))

# create an individual variable for each 12 hour period, up to 120 hours, fill with 1 if IMV or 0 if CPAP 

tbl1 <- 
  tbl1 %>% 
  mutate("X_at_12" = if_else(cpap > 12, 1, 0)) %>% 
  mutate("X_at_24" = if_else(cpap > 24, 1, 0)) %>% 
  mutate("X_at_36" = if_else(cpap > 36, 1, 0)) %>% 
  mutate("X_at_48" = if_else(cpap > 48, 1, 0)) %>% 
  mutate("X_at_60" = if_else(cpap > 60, 1, 0))

# create a random nudge at each 12 hour window:

tbl1 <- 
  tbl1 %>% 
  mutate("Z_at_12" = sample(0:1, size = 100, replace = T)) %>% 
  mutate("Z_at_24" = sample(0:1, size = 100, replace = T)) %>% 
  mutate("Z_at_36" = sample(0:1, size = 100, replace = T)) %>% 
  mutate("Z_at_48" = sample(0:1, size = 100, replace = T)) %>% 
  mutate("Z_at_60" = sample(0:1, size = 100, replace = T))

```



```{r}

# actually what we want to do is START by generating the random nudges, then extrapolate whether they were followed or not based on a compliance rate of x% and then obtain the duration of cpap, and then look at los (at which point we could derive a formula to force a correlation between los and cpap duration to enable the rest of the analysis to run).

# first do it for 2 shifts (24 hours)

tbl2 <- 
  tbl1 %>% 
  select("n","X_at_12", "Z_at_12", "X_at_24", "Z_at_24")

# the patient state at 12 hours should first be observed (X_at_12), then the nudge randomisation result (Z_at_12), then the next X value should indicate whether the allocation was complied with or not.

# change the first X allocation to a random sample rather than derived based on duration of cpap as in the first attempt

tbl2 <- 
  tbl2 %>% 
  mutate(X_at_12 = sample(0:1, size = 100, replace = T))

# if the state of X at 12 hours is 1 (IMV) then no further randomisation occurs and no further monitoring is required so the rest of the columns should be NA.

# do this stepwise:

# 1.  Create the state of the patient (X) at the 12 hour review, then assign a value as to whether or not they are eligible for randomisation at that point (T|F), if they are ventilated, then they will not be randomised.

tbl2 <- 
  tbl2 %>% 
  mutate(elig_for_R = if_else(X_at_12 == 1, FALSE, TRUE)) %>% 
  select("n","X_at_12", "elig_for_R")

# 2. create an empty column of NAs for the randomisation allocation at 12 hours:

tbl2 <- 
  tbl2 %>% 
  mutate("Z_at_12" = NA)

# 3. highlight the patients eligible for randomisation and leave the rest with NA values

tbl2 <- 
  tbl2 %>% 
  mutate(Z_at_12 = ifelse(elig_for_R == T, T, NA))

# 4. replace the T value with a randomly generated 0 or 1 for only those eligible:

tbl2 <- 
  tbl2 %>% 
  mutate(Z_at_12 = ifelse(Z_at_12 == T, sample(0:1, 
                                               size = sum(!is.na(Z_at_12)), 
                                               replace = T)))


# after determining eligibility for re-randomisation, the clinician receives the nudge prompt to consider re-randomisation:

# "Do you have equipoise to submit this patient for randomisation?"

# the clinician can either say YES or NO at this point.

# start with the assumption that in 10% of cases the clinician will have enough equipoise to submit the patient to re-randomisation.  

# so for 10% of those without NA values at Z_at_12, the clinician would "unmask" the randomisation allocation:

  # select the eligible patients into a new object:

elig_sample <- 
  tbl2 %>% 
  filter(!is.na(Z_at_12))

  # select a random sample of 10% of the eligible patients who the clinician has equipoise for:

ten_percent <- 
  elig_sample %>% 
  sample_frac(0.10)

  # create a new column for the selected, eligible patients in the main table:

tbl2 <- 
  tbl2 %>% 
  mutate(equipoise_at_12 = if_else(tbl2$n %in% ten_percent$n, T, F))

# NB - you could do this the othe way around so the patients are randomised at this point, rather than just unmasked, I don't know if this would make a difference or not?

  # it feels like this makes more sense, so the eligible people would be subject to a 1:1 randomisation between cpap and imv.

# the next decision is, for those eligible patients, with equipoise, does the clinician then comply with the unmasked allocation recommendation:

# 
 

# next we observe the patient state (X) at 24 hours (the next randomisation screening point).

# do this in two stages:
  # A) an indication of whether the clinician complied with the nudge or not

  # B) an indication of whether there was unintended cross over from cpap to imv during the following 12 hour period


# A.  Nudge compliance:

# the possible options upon receipt of the first nudge randomisation are:

  # 1. Compliance with the nudge (10%)

  # 2. Non-compliance with the nudge (90%)

# we could apply estimated values to both parts e.g. 10% compliance.

# compliance with the nudge is not one-directional.  i.e. we might allocate a 10% compliance with the nudge, but it may be more if cpap and less if IMV depending on preferences.

# for simplicity now, keep it as a single value for both treatment arms.  

# as a starting point we will assume that in 90% of the eligible patients the clinician does not comply with the treatment recommendation.  

# 3. Unintentional, one-way cross over from the cpap group to the IMV group due to patient deterioration.

# apply estimated values for the above quantities:

  # 1. 10% i.e. for 1 in 10 patients randomised, the 





```






# Bibliography

1. [@walker2017]
2. [@keele2018]
